Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-06-24 ms.tex
Adding paragraph discussing x_s plot

The Monte Carlo method is used to solve the transfer equation numerically as a point of reference for these analytic solutions. Photon packets are initialized at a monochromatic source frequency $x_s$ and are allowed to propagate through the medium until escaping, at which point their positions, outgoing angles, and frequencies are logged. Frequency redistribution is calculated at each scattering, including the effects of recoil, to obtain the spectrum at the surface of the spherical simulation domain. Further details of the Monte Carlo simulation are discussed in \cite{2017ApJ...851..150H}.

@@

The relative size of the boundary term $H_{\rm bc}$ with optical depth is shown graphically in Fig. \ref{fig:taubc}, where $H_{\rm bc}$ is shown for several optical depths along with Monte Carlo data. Since the Monte Carlo simulation represents the true solution, it can be compared directly against $H_{\rm bc}$ by subtracting off the homogeneous piece of the solution ($H_{\rm bc} = H - H_0$). From Eq. \ref{eq:hbc_scaling}, the size of the frequency-dependent boundary condition correction should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. This factor has been scaled out of the figure such that solutions for different optical depths should show close agreement on the figure if the relation holds. Indeed, agreement between $H_{\rm bc}$ and the Monte Carl points improves as the optical depth becomes larger. At lower optical depths, the scattering of photons within the Doppler core of the line becomes important, yet our solution is not constructed to model this behavior precisely. The effects of core scattering can be seen in the Monte Carlo data, however, as the peaks of the distribution shift outward in frequency from the line core around $\tau_0 = 10^5$. Thus, the $(a\tau)^{1/3}$ scaling holds only for larger optical depths where scattering in the core can be ignored.

@@

We now compare each of the solutions for surface flux with Monte Carlo simulations for a limited range of physical parameters $\tau_0$ and $x_s$. In Fig. \ref{fig:sol_mc_residual_0}, the Monte Carlo is shown with solutions $H_{\rm d}$, $H_{\rm 0}$, and $H_{\rm 0 + bc} = H_{\rm 0} + H_{\rm bc}$ overplotted for an optical depth of $\tau_0 = 10^7$ and photons emitted at line center $\sigma_s = \sigma_0$. The solution with the correct frequency-dependent boundary condition enforced, $H_{\rm 0 + bc}$, has considerably lower residuals to Monte Carlo than the other solutions, especially in the line wing. It is evident that the divergent solution $H_{\rm d}$ fails in the line wings. The $H_{\rm bc}$ term, when added to the solution $H_0$ derived by \citet{2006ApJ...649...14D}, adequately corrects for the apparent excess of flux in the peaks of the spectrum. This peak excess is also present in previous work that used the \citet{1990ApJ...350..216N} solution, such as \citet{2015MNRAS.449.4336S}. The boundary term resolves the slight deficit of $H_{\rm0}$ in the line wings, further improving agreement with the numerical result.

@@

>>>>

The Monte Carlo method is used to solve the transfer equation numerically as a point of reference for these analytic solutions. For each simulation, a total of $2^20 \approx 10^6$ photon packets are initialized at a monochromatic source frequency $x_s$ and are allowed to propagate through the medium until escaping, at which point their positions, outgoing angles, and frequencies are logged. A constant temperature of $T=10000\ \rm K$ is set for the gas. Frequency redistribution is calculated at each scattering, including the effects of recoil, to obtain the spectrum at the surface of the spherical simulation domain. Further details of the Monte Carlo simulation are discussed in \cite{2017ApJ...851..150H}.

@@

The relative size of the boundary term $H_{\rm bc}$ with optical depth is shown graphically in Fig. \ref{fig:taubc}, where $H_{\rm bc}$ is shown for several optical depths along with Monte Carlo data. Since the Monte Carlo simulation represents the true solution, it can be compared directly against $H_{\rm bc}$ by subtracting off the homogeneous piece of the solution ($H_{\rm bc} = H - H_0$). From Eq. \ref{eq:hbc_scaling}, the size of the frequency-dependent boundary condition correction should become smaller with larger optical depths, following a $(a\tau_0)^{-1/3}$ scaling. This factor has been scaled out of the figure such that solutions for different optical depths should show close agreement if the relation holds. Indeed, agreement between $H_{\rm bc}$ and the Monte Carl points improves as the optical depth becomes larger. At lower optical depths, the scattering of photons within the Doppler core of the line becomes important, yet our solution is not constructed to model this behavior precisely. The effects of core scattering can be seen in the Monte Carlo data, however, as the peaks of the distribution shift outward in frequency from the line core around $\tau_0 = 10^5$. Thus, the $(a\tau)^{1/3}$ scaling holds only for larger optical depths where scattering in the core can be ignored.

@@

We now compare each of the solutions for surface flux with Monte Carlo simulations. In Fig. \ref{fig:sol_mc_residual_0}, the Monte Carlo is shown with solutions $H_{\rm d}$, $H_{\rm 0}$, and $H_{\rm 0 + bc} = H_{\rm 0} + H_{\rm bc}$ over-plotted for an optical depth of $\tau_0 = 10^7$ and photons emitted at line center $\rm x_s = 0$. The solution with the correct frequency-dependent boundary condition enforced, $H_{\rm 0 + bc}$, has considerably lower residuals to Monte Carlo than the other solutions, especially in the line wing. It is evident that the divergent solution $H_{\rm d}$ fails in the line wings. The $H_{\rm bc}$ term, when added to the solution $H_0$ derived by \citet{2006ApJ...649...14D}, adequately corrects for the apparent excess of flux in the peaks of the spectrum. This peak excess is also present in previous work that used the \citet{1990ApJ...350..216N} solution, such as \citet{2015MNRAS.449.4336S}. The boundary term resolves the slight deficit of $H_{\rm0}$ in the line wings, further improving agreement with the numerical result. Note that as in Fig. \ref{fig:taubc} the residuals to the $H_0$ solution are a close match to the $H_{\rm bc}$ term.

@@

Since the solutions have been generalized to allow for a monochromatic source of photons at a frequency $\sigma_s$ away from line center, their agreement with Monte Carlo as a function of source frequency can also be characterized. Photons initialized closer to the line wing may travel larger mean free paths as the opacity is sensitively dependent on the line profile. The larger spatial diffusion contributes to greater escape probability, biasing the frequency distribution of escaping photons. In the limit that $|\rm x_s|$ becomes large, the distribution becomes a delta function in frequency as all photons escape the gas without scattering. Fig. \ref{fig:sol_mc_xinit} shows calculations performed for three initial photon frequencies $\rm x_s = 0.0, 6.0$, and $12.0$ for a line center optical depth of $\tau_0=10^7$. The bias in frequency distribution is slight for $\rm x_s = 6.0$ where the line profile is still relatively large, but the increase in spatial diffusion is much more substantial when photons are initialized with $\rm x_s=12.0$, further away from the line core. On this point, we note that the discrepancy between the Monte Carlo data and $H_0$ (effectively the scale of the corrective $H_{\rm bc}$ term) becomes larger as $\rm x_s$ increases. Thus, for sufficiently large optical depths and emission away from line center, the importance of enforcing frequency-dependent boundary conditions at the surface of the sphere grows. %Note: should we characterize this scaling quantitatively?
